\begin{center} \bf \sffamily \LARGE Prisoner's dilemma with stochastic payoffs\\[-1cm] \end{center}

### 2023-02-21 main.tex
working on paper in folder paper

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.70\textwidth]{static/expected_and_stochastic_for_donation_game.pdf}
    % \includegraphics[width=.70\textwidth]{static/expected_and_stochastic_for_donation_game_b_10.pdf}
    \caption{{\bf Evolutionary dynamics under expected payoffs and last round with one interaction payoffs.} 
    We have run two simulations of the evolutionary process described in
    section~\ref{section:model} for $t\!=\!10^7$ time steps. For each time step,
    we have recorded the current resident population ($y,p,q$). Since
    simulations are run for a relatively high continuation probability of
    $\delta\!=\!0.999$, we do not report the players' initial cooperation
    probability $y$. The graphs show how often the resident population chooses
    each combination ($p,q$) of conditional cooperation probabilities in the
    subsequent rounds. ({\bf A}) If players update based on their expected
    payoffs, the resident population typically applies a strategy for which
    $p\!\approx\!1$ and $q\!\le\!1\!-\!c/b\!=\!0.9$. ({\bf B}) When players
    update their strategies based on their realized payoffs in the last round,
    there are two different predominant behaviors. The resident population
    either consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of
    conditional cooperators. In the latter case, the maximum level of $q$
    consistent with stable cooperation is somewhat smaller compared to the
    expected-payoff setting, $q\!<\!0.5$. The cooperation rate within the
    resident population (averaged over all games and over all time steps) is
    close to 100\%. Parameters: $N\!=\!100$, $c\!=\!1$, $\beta\!=\!1$,
    $\delta\!=\!0.999$.}
    \label{fig:expected_and_stochastic_for_donation}
\end{figure}

@@

We also investigate the effect of the strength of selection $\beta$.
Figure~\ref{fig:cooperation_rate_over_betas} illustrates results for various
runs of the evolutionary process. For weak selection, \(\beta < 1\), we observe
that the two methods yield similar results, however, as \(\beta\) increases
there is variation in the evolving populations. In the case of expected payoffs
the resident populations become more cooperative as \(\beta\) increases, whereas
in the case of limited memory payoffs, the resident populations become more
defective.

\begin{figure}[!htbp]
  \centering
    \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
\caption{{\bf The evolution of cooperation for different selection strength values.}
We vary the selection strength $\beta$. In all cases, stochastic payoff
evaluation tends to reduce the evolving cooperation rates. ({\bf A}) the
probabilities \(p, q\) for resident population over \(10^7\) time steps for each
\(\beta\) value. ({\bf B}) The cooperation rate within the resident population
(averaged over all games and over all time steps) over \(\beta\). Unless
explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$,
$c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
$t\!=\!5\times 10^7$ time steps for each parameter
combination.}\label{fig:cooperation_rate_over_betas}
\end{figure}

\begin{figure}[!htbp]
  \centering
    \includegraphics[width=\textwidth]{static/more_memory.pdf}
% \caption{{\bf The evolution of cooperation for different selection strength values.}
% We vary the selection strength $\beta$. In all cases, stochastic payoff
% evaluation tends to reduce the evolving cooperation rates. ({\bf A}) the
% probabilities \(p, q\) for resident population over \(10^7\) time steps for each
% \(\beta\) value. ({\bf B}) The cooperation rate within the resident population
% (averaged over all games and over all time steps) over \(\beta\). Unless
% explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$,
% $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
% $t\!=\!5\times 10^7$ time steps for each parameter
% combination.}\label{fig:cooperation_rate_over_betas}
\end{figure}

% \subsection{Effect of updating payoffs in different social
% dilemmas}\label{section:2_by_2_games}

% In the previous section we gained insights into the effects of the updating
% payoffs, and into how parameters such as the benefit and the strength of
% selection can intensify them. We investigated these effects by using the
% donation game. In order to broaden our understanding of the updating payoffs on
% different forms of possible human interactions we extend our approach to all \(2
% \times 2\) symmetric games. More specifically, we apply our analysis to four
% different classes of games; the harmony, the snowdrift, the prisoner's dilemma
% and the stag-hunt games.

% We compare the results of the evolutionary process when the updating payoffs are
% based on the expected payoffs, and on the last round payoffs for all the four
% possible classes of games (Figures~\ref{fig:expected_two_by_two}
% and~\ref{fig:last_round_two_by_two}). In Figure~\ref{fig:expected_two_by_two}
% each sub figure represents a run of evolutionary process for a different set of
% values for \(\mathcal{U}\). Without loss of generality we set \(R=1\) and
% \(P=0\)~\cite{Martinez2012, Roca2009}, and we vary the values of the temptation
% payoff \(T\) (across the \(x\)-axis) and of the sucker's payoff \(S\) (across
% the \(y\)-axis). Starting at the upper left corner and proceeding clockwise the
% quadrants correspond to the harmony, the snowdrift, the prisoner's dilemma and
% the stag-hunt games.

% \begin{figure}[!htbp]
%   \centering
%   \includegraphics[width=\textwidth]{static/expected_two_by_two_games.pdf}
%   \caption{{\bf Evolutionary dynamics under expected payoffs for various social dilemmas.}
%   We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
%   the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
%   \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
%   corresponds to the harmony game. The preference ordering for the harmony game
%   is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T
%   > 1\) corresponds to the snowdrift game. The preference ordering for the
%   snowdrift game is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where
%   \(S < 0\) and \(T \leq 1\) corresponds to the stag-hunt game. The preference
%   ordering for the stag-hunt game is \(R > T > P > S\). ({\bf D}) The bottom
%   right quadrant where \(S < 0\) and \(T > 1\) corresponds to the prisoner's
%   dilemma game. The preference ordering for the prisoner's dilemma game is \(T >
%   R > P > S\). Unless explicitly varied, the parameters of the simulation are
%   $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
%   \label{fig:expected_two_by_two}
% \end{figure}

% \begin{figure}[!htbp]
%   \centering
%   \includegraphics[width=\textwidth]{static/stochastic_two_by_two_games.pdf}
%   \caption{{\bf Evolutionary dynamics under last round payoffs for various social dilemmas.} 
%   We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
%   the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
%   \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
%   corresponds to the harmony game. The preference ordering for the harmony game
%   is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T
%   > 1\) corresponds to the snowdrift game. The preference ordering for the
%   snowdrift game is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where
%   \(S < 0\) and \(T \leq 1\) corresponds to the stag-hunt game. The preference
%   ordering for the stag-hunt game is \(R > T > P > S\). ({\bf D}) The bottom
%   right quadrant where \(S < 0\) and \(T > 1\) corresponds to the prisoner's
%   dilemma game. The preference ordering for the prisoner's dilemma game is \(T >
%   R > P > S\). Unless explicitly varied, the parameters of the simulation are
%   $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
%   \label{fig:last_round_two_by_two}
% \end{figure}

% The harmony game describes an interaction without conflict. In the harmony game
% it is in the best interest for both players to cooperate, and in an evolutionary
% process cooperators prevail. We observe that for most harmony game runs the
% resident population overwhelmingly applies a strategy for which \(p\) and \(q\)
% are \(\approx 1\) (Figure~\ref{fig:expected_two_by_two}). The snowdrift game
% describes a situation similar to that of the prisoner's dilemma; cooperation
% results in a benefit to the opposing player but entailing a cost to the
% cooperator. However, in the snowdrift game individuals obtain immediate direct
% benefits from the cooperative acts which leads to \(S > P\). In the snowdrift
% game more cooperation emerges compared to the prisoner's dilemma. Defection is
% never a resident strategy and for the same values of temptation the overall
% \(q-\) values are higher. The last class of games we present are for the
% stag-hunt game. In the stag-hunt game there is an equilibrium in which both
% players cooperate as well as one in which both defect. For small values of the
% temptation and the sucker's payoffs we observe that cooperation prevails similar
% to the harmony game. However, in the case where the temptation and the reward
% payoffs are equal the resident population either consists of defectors or
% cooperators.

% Although the patterns of the evolved populations remain similar if we consider
% that players update their strategies based on their last round payoff
% (Figure~\ref{fig:last_round_two_by_two}), there are still dissimilarities
% between the two approaches. To better understand these dissimilarities, we
% explore the cooperation rates for each of the cases of the social dilemmas we
% have presented (Figure~\ref{fig:cooperation_rates_two_by_two_part_one}).

% In the harmony game the cooperation rates for both approaches are similar. The
% biggest difference occurs for \(T=1\), nevertheless, it is less than 10\%. There
% are two instances for which the cooperation rates are higher for the limited
% memory payoffs compared to the expected. These are for \(T=2.2, S=0.4\) and for
% \(T=2.2, S=0.8\). It is only in the snowdrift game that cooperation is
% overestimated when a minimum of information is considered. Players are slightly
% more likely to cooperate after a cooperation and to forgive after being at the
% receiving end of a defection. For the rest of the cases the classical scenario
% overestimates the average cooperation rate, supporting the results of the
% previous section. There is a substantial difference in the cooperation rates for
% the prisoner's dilemma class. We observe that for \(S<-0.4\) cooperation almost
% never emerges if players imitate based only on the last round payoff.

% \begin{figure}[!htbp]
%   \centering
%   \includegraphics[width=.4\textwidth]{static/expected_two_by_two_games_cooperation.pdf}
%   \includegraphics[width=.4\textwidth]{static/stochastic_two_by_two_games_cooperation.pdf}
%   \caption{\textbf{Cooperation rates for various social dilemmas.} ({\bf A}) If
%   players update based on their expected payoffs. ({\bf B}) If players update
%   based on their last round payoffs. We vary the temptation payoff \(T \in
%   \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across the \(x\) axis, and \(S \in \{1.2,
%   0.8, 0.4, 0, -0.4, -0.8\}\) across the \(y\) axis. Unless explicitly varied,
%   the parameters of the simulation are $N\!=\!100$, $\beta\!=\!10$,
%   $\delta\!=\!0.99$, $t\!=\!5\times
%   10^7$.}\label{fig:cooperation_rates_two_by_two_part_one}
% \end{figure}

% So far we have explored the difference between the expected payoffs and the last
% round payoffs. In order to explore further the effect of limited memory we allow
% individuals to remember more. More precisely, up to two interactions and up to
% the last two rounds. In total we present results for three more updating
% payoffs. These are the payoffs when individuals consider the last two rounds
% with another member of the population, the last round with two members of the
% population, and the last two rounds with two members of the population. Similar
% to the last round payoff, we use simulations and record which strategies the
% players adopt over time based
% (Appendix~\ref{appendix:further_simulation_results}) and we compare the evolving
% cooperation rates to those of the expected payoffs
% (Figure~\ref{fig:cooperation_rates_two_by_two_part_two}).

% In the case of the last two rounds payoff we make two observations. Initially,
% for the harmony, the stag-hunt and the prisoner's dilemma classes the average
% cooperation rates are higher than the previous case of the last round payoff.
% The rates remain strictly less than in the expected payoffs, however, in the
% case of the prisoner's dilemma we observe a significant increase. Secondly, for
% the snowdrift class there are more instances for which the cooperation rates are
% higher than in the expected payoffs. This could indicate that in the case of the
% snowdrift game the expected payoffs underestimate cooperation. Once players
% interact with two other members of the population, instead of one, the expected
% payoffs strictly yield a higher cooperation rate  even in the case of the
% snowdrift game. Similar to the previous result, the difference between the two
% approaches is always higher for the prisoner's dilemma class.

% \begin{figure}[!htbp]
%   \centering
%   \includegraphics[width=.33\textwidth]{static/rounds_two_by_two_games_cooperation.pdf}
%   \includegraphics[width=.33\textwidth]{static/opponents_two_by_two_games_cooperation.pdf}
%   \includegraphics[width=.33\textwidth]{static/rounds_opponents_two_by_two_games_cooperation.pdf}
%   \caption{\textbf{Cooperation rates over various social dilemmas for different
%   limited memory payoffs.} ({\bf A}) If players update based on their last two
%   rounds with another member of the population. ({\bf B}) If players update
%   based on their last round with two other members of the population. ({\bf C})
%   If players update based on their last two rounds with two other members of the
%   population. We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8,
%   2.2\}\) across the \(x\) axis, and \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
%   across the \(y\) axis. Unless explicitly varied, the parameters of the
%   simulation are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times
%   10^7$..}
%   \label{fig:cooperation_rates_two_by_two_part_two}
% \end{figure}

\section{Conclusions}\label{section:conclusions}

Cooperation can be seen at odd, why is it that we choose to help others,
increasing their payoff, at the expense of decreasing one's own? In spite of all
the selfish genes’ animal and human communities seem to altruistically help each
other and cooperate, and evolutionary game theory has helped us shape our
understanding of the evolution of cooperation.

Previous evolutionary models often feature a curious inconsistency. When
modeling how individuals make decisions in each round, these models assume that
players only remember the last round. However, when modeling how individuals
update their strategies over time, individuals are assumed to have perfect
memory.

Here, we have explored how robust cooperation is as models deviate from the
perfect memory assumption. Initially we considered the donation game. We showed
that when the last round payoff is used instead of the expected payoffs,
cooperation can even evolve when individuals only use a minimum of information,
however, the evolving cooperation rates are typically lower. The resident
strategies were both less cooperative and less generous. This effect was only
intensified we increase the benefit and the strength of selection independently.
The results showed that as each parameter was being increased the difference
between the cooperation rates were widening. This indicates that cooperative
players benefit from being able to interact with everyone in the population.

We extended our approach and presented results not only based on different
classes of social games, but also based on different limited memory payoffs. The
analysis showed that specifically for the case of the prisoner's dilemma games
the contrast between the two approaches can be significantly large. The
prisoner's dilemma is one of the most applied types of social games when we
discuss the evolution of cooperation. Our results indicate that cooperation
struggles to evolve in the prisoner's dilemma when only a minimum of social
information is used at the updating stage. Interestingly our simulations also
showed that in same cases cooperation can benefit from limited memory. This was
specific for the snowdrift game.

\section{Acknowledgements}

This work was supported by the European Research Council Starting Grant 850529:
E-DIRECT.

\appendix

\section{Model Setup}\label{appendix:methods}

Consider a population of \(N\) individuals where \(N\) is even. At any point in
time there are at most two different strategies in present in the population.
More specifically, a mutant strategy played by \(k\) individuals and a resident
strategy played by \(N - k\) individuals. We assume a pairwise process in which
strategies spread because they are imitated more often. Each step of the
evolutionary process consists of two stages; a game stage and an update stage.

In the game stage, each individual is randomly matched with some other
individual in the population. Their interaction lasts for a number of turns
which is not fixed but depends on the continuation probability \(\delta\). At
each turn the individuals choose between cooperation (\(C\)) and defection
(\(D\)). Thus, there are four possible outcomes in each turn \(CC, CD, DC\) and
\(DD\). If both players cooperate they receive the reward payoff \(R\), whereas
if both players defect they receive the punishment payoff \(P\). If one
cooperates but the other defects, the defector receives the temptation to
defect, \(T\), whereas the cooperator receives the sucker's payoff, \(S\). Let
$\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round, and
let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector. We present
results for various values of $\mathcal{U}$ for all the symmetric \(2 \times 2\)
games.

A further assumption of our model is that individuals make use of reactive
strategies when they make decisions in each round. Reactive strategies are a set
of strategies that take into account only the previous action of the opponent. A
reactive strategy can be written explicitly as a vector,

\[s=(y, p, q)\]

where \(y\) is the probability that the strategy opens with a cooperation and
\(p, q\) are the probabilities that the strategy cooperates given that the
opponent cooperated and defected equivalently.

In the updating stage, two players are randomly drawn from the population, a
`learner' and a `role model'. The learner adopts the role model's strategy based
on the Fermi distribution function, %ToDo add reference

where $u_{L}\!\in\! \mathcal{U}$ is the learner's payoff, $u_{RM}\!\in\!
\mathcal{U}$ is the role model's payoff, and $\beta\!\ge\!0$ is the strength of
selection.

>>>>

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=1.5\textwidth]{static/donation_expected_last_round_summary_results.pdf}
    % \includegraphics[width=.70\textwidth]{static/expected_and_stochastic_for_donation_game_b_10.pdf}
    \caption{{\bf Evolutionary dynamics under expected payoffs and last round with one interaction payoffs.} 
    We have run two simulations of the evolutionary process described in
    section~\ref{section:model} for $t\!=\!10^7$ time steps. For each time step,
    we have recorded the current resident population ($y,p,q$). Since
    simulations are run for a relatively high continuation probability of
    $\delta\!=\!0.999$, we do not report the players' initial cooperation
    probability $y$. The graphs show how often the resident population chooses
    each combination ($p,q$) of conditional cooperation probabilities in the
    subsequent rounds. ({\bf A}) If players update based on their expected
    payoffs, the resident population typically applies a strategy for which
    $p\!\approx\!1$ and $q\!\le\!1\!-\!c/b\!=\!0.9$. ({\bf B}) When players
    update their strategies based on their realized payoffs in the last round,
    there are two different predominant behaviors. The resident population
    either consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of
    conditional cooperators. In the latter case, the maximum level of $q$
    consistent with stable cooperation is somewhat smaller compared to the
    expected-payoff setting, $q\!<\!0.5$. The cooperation rate within the
    resident population (averaged over all games and over all time steps) is
    close to 100\%. Parameters: $N\!=\!100$, $c\!=\!1$, $\beta\!=\!1$,
    $\delta\!=\!0.999$.}
    \label{fig:expected_and_stochastic_for_donation}
\end{figure}

@@

% We also investigate the effect of the strength of selection $\beta$.
% Figure~\ref{fig:cooperation_rate_over_betas} illustrates results for various
% runs of the evolutionary process. For weak selection, \(\beta < 1\), we observe
% that the two methods yield similar results, however, as \(\beta\) increases
% there is variation in the evolving populations. In the case of expected payoffs
% the resident populations become more cooperative as \(\beta\) increases, whereas
% in the case of limited memory payoffs, the resident populations become more
% defective.

% \begin{figure}[!htbp]
%   \centering
%     \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
% \caption{{\bf The evolution of cooperation for different selection strength values.}
% We vary the selection strength $\beta$. In all cases, stochastic payoff
% evaluation tends to reduce the evolving cooperation rates. ({\bf A}) the
% probabilities \(p, q\) for resident population over \(10^7\) time steps for each
% \(\beta\) value. ({\bf B}) The cooperation rate within the resident population
% (averaged over all games and over all time steps) over \(\beta\). Unless
% explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$,
% $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
% $t\!=\!5\times 10^7$ time steps for each parameter
% combination.}\label{fig:cooperation_rate_over_betas}
% \end{figure}

% \begin{figure}[!htbp]
%   \centering
%     \includegraphics[width=\textwidth]{static/more_memory.pdf}
% % \caption{{\bf The evolution of cooperation for different selection strength values.}
% % We vary the selection strength $\beta$. In all cases, stochastic payoff
% % evaluation tends to reduce the evolving cooperation rates. ({\bf A}) the
% % probabilities \(p, q\) for resident population over \(10^7\) time steps for each
% % \(\beta\) value. ({\bf B}) The cooperation rate within the resident population
% % (averaged over all games and over all time steps) over \(\beta\). Unless
% % explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$,
% % $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
% % $t\!=\!5\times 10^7$ time steps for each parameter
% % combination.}\label{fig:cooperation_rate_over_betas}
% \end{figure}

% % \subsection{Effect of updating payoffs in different social
% % dilemmas}\label{section:2_by_2_games}

% % In the previous section we gained insights into the effects of the updating
% % payoffs, and into how parameters such as the benefit and the strength of
% % selection can intensify them. We investigated these effects by using the
% % donation game. In order to broaden our understanding of the updating payoffs on
% % different forms of possible human interactions we extend our approach to all \(2
% % \times 2\) symmetric games. More specifically, we apply our analysis to four
% % different classes of games; the harmony, the snowdrift, the prisoner's dilemma
% % and the stag-hunt games.

% % We compare the results of the evolutionary process when the updating payoffs are
% % based on the expected payoffs, and on the last round payoffs for all the four
% % possible classes of games (Figures~\ref{fig:expected_two_by_two}
% % and~\ref{fig:last_round_two_by_two}). In Figure~\ref{fig:expected_two_by_two}
% % each sub figure represents a run of evolutionary process for a different set of
% % values for \(\mathcal{U}\). Without loss of generality we set \(R=1\) and
% % \(P=0\)~\cite{Martinez2012, Roca2009}, and we vary the values of the temptation
% % payoff \(T\) (across the \(x\)-axis) and of the sucker's payoff \(S\) (across
% % the \(y\)-axis). Starting at the upper left corner and proceeding clockwise the
% % quadrants correspond to the harmony, the snowdrift, the prisoner's dilemma and
% % the stag-hunt games.

% % \begin{figure}[!htbp]
% %   \centering
% %   \includegraphics[width=\textwidth]{static/expected_two_by_two_games.pdf}
% %   \caption{{\bf Evolutionary dynamics under expected payoffs for various social dilemmas.}
% %   We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
% %   the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
% %   \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
% %   corresponds to the harmony game. The preference ordering for the harmony game
% %   is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T
% %   > 1\) corresponds to the snowdrift game. The preference ordering for the
% %   snowdrift game is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where
% %   \(S < 0\) and \(T \leq 1\) corresponds to the stag-hunt game. The preference
% %   ordering for the stag-hunt game is \(R > T > P > S\). ({\bf D}) The bottom
% %   right quadrant where \(S < 0\) and \(T > 1\) corresponds to the prisoner's
% %   dilemma game. The preference ordering for the prisoner's dilemma game is \(T >
% %   R > P > S\). Unless explicitly varied, the parameters of the simulation are
% %   $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
% %   \label{fig:expected_two_by_two}
% % \end{figure}

% % \begin{figure}[!htbp]
% %   \centering
% %   \includegraphics[width=\textwidth]{static/stochastic_two_by_two_games.pdf}
% %   \caption{{\bf Evolutionary dynamics under last round payoffs for various social dilemmas.} 
% %   We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
% %   the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
% %   \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
% %   corresponds to the harmony game. The preference ordering for the harmony game
% %   is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T
% %   > 1\) corresponds to the snowdrift game. The preference ordering for the
% %   snowdrift game is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where
% %   \(S < 0\) and \(T \leq 1\) corresponds to the stag-hunt game. The preference
% %   ordering for the stag-hunt game is \(R > T > P > S\). ({\bf D}) The bottom
% %   right quadrant where \(S < 0\) and \(T > 1\) corresponds to the prisoner's
% %   dilemma game. The preference ordering for the prisoner's dilemma game is \(T >
% %   R > P > S\). Unless explicitly varied, the parameters of the simulation are
% %   $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
% %   \label{fig:last_round_two_by_two}
% % \end{figure}

% % The harmony game describes an interaction without conflict. In the harmony game
% % it is in the best interest for both players to cooperate, and in an evolutionary
% % process cooperators prevail. We observe that for most harmony game runs the
% % resident population overwhelmingly applies a strategy for which \(p\) and \(q\)
% % are \(\approx 1\) (Figure~\ref{fig:expected_two_by_two}). The snowdrift game
% % describes a situation similar to that of the prisoner's dilemma; cooperation
% % results in a benefit to the opposing player but entailing a cost to the
% % cooperator. However, in the snowdrift game individuals obtain immediate direct
% % benefits from the cooperative acts which leads to \(S > P\). In the snowdrift
% % game more cooperation emerges compared to the prisoner's dilemma. Defection is
% % never a resident strategy and for the same values of temptation the overall
% % \(q-\) values are higher. The last class of games we present are for the
% % stag-hunt game. In the stag-hunt game there is an equilibrium in which both
% % players cooperate as well as one in which both defect. For small values of the
% % temptation and the sucker's payoffs we observe that cooperation prevails similar
% % to the harmony game. However, in the case where the temptation and the reward
% % payoffs are equal the resident population either consists of defectors or
% % cooperators.

% % Although the patterns of the evolved populations remain similar if we consider
% % that players update their strategies based on their last round payoff
% % (Figure~\ref{fig:last_round_two_by_two}), there are still dissimilarities
% % between the two approaches. To better understand these dissimilarities, we
% % explore the cooperation rates for each of the cases of the social dilemmas we
% % have presented (Figure~\ref{fig:cooperation_rates_two_by_two_part_one}).

% % In the harmony game the cooperation rates for both approaches are similar. The
% % biggest difference occurs for \(T=1\), nevertheless, it is less than 10\%. There
% % are two instances for which the cooperation rates are higher for the limited
% % memory payoffs compared to the expected. These are for \(T=2.2, S=0.4\) and for
% % \(T=2.2, S=0.8\). It is only in the snowdrift game that cooperation is
% % overestimated when a minimum of information is considered. Players are slightly
% % more likely to cooperate after a cooperation and to forgive after being at the
% % receiving end of a defection. For the rest of the cases the classical scenario
% % overestimates the average cooperation rate, supporting the results of the
% % previous section. There is a substantial difference in the cooperation rates for
% % the prisoner's dilemma class. We observe that for \(S<-0.4\) cooperation almost
% % never emerges if players imitate based only on the last round payoff.

% % \begin{figure}[!htbp]
% %   \centering
% %   \includegraphics[width=.4\textwidth]{static/expected_two_by_two_games_cooperation.pdf}
% %   \includegraphics[width=.4\textwidth]{static/stochastic_two_by_two_games_cooperation.pdf}
% %   \caption{\textbf{Cooperation rates for various social dilemmas.} ({\bf A}) If
% %   players update based on their expected payoffs. ({\bf B}) If players update
% %   based on their last round payoffs. We vary the temptation payoff \(T \in
% %   \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across the \(x\) axis, and \(S \in \{1.2,
% %   0.8, 0.4, 0, -0.4, -0.8\}\) across the \(y\) axis. Unless explicitly varied,
% %   the parameters of the simulation are $N\!=\!100$, $\beta\!=\!10$,
% %   $\delta\!=\!0.99$, $t\!=\!5\times
% %   10^7$.}\label{fig:cooperation_rates_two_by_two_part_one}
% % \end{figure}

% % So far we have explored the difference between the expected payoffs and the last
% % round payoffs. In order to explore further the effect of limited memory we allow
% % individuals to remember more. More precisely, up to two interactions and up to
% % the last two rounds. In total we present results for three more updating
% % payoffs. These are the payoffs when individuals consider the last two rounds
% % with another member of the population, the last round with two members of the
% % population, and the last two rounds with two members of the population. Similar
% % to the last round payoff, we use simulations and record which strategies the
% % players adopt over time based
% % (Appendix~\ref{appendix:further_simulation_results}) and we compare the evolving
% % cooperation rates to those of the expected payoffs
% % (Figure~\ref{fig:cooperation_rates_two_by_two_part_two}).

% % In the case of the last two rounds payoff we make two observations. Initially,
% % for the harmony, the stag-hunt and the prisoner's dilemma classes the average
% % cooperation rates are higher than the previous case of the last round payoff.
% % The rates remain strictly less than in the expected payoffs, however, in the
% % case of the prisoner's dilemma we observe a significant increase. Secondly, for
% % the snowdrift class there are more instances for which the cooperation rates are
% % higher than in the expected payoffs. This could indicate that in the case of the
% % snowdrift game the expected payoffs underestimate cooperation. Once players
% % interact with two other members of the population, instead of one, the expected
% % payoffs strictly yield a higher cooperation rate  even in the case of the
% % snowdrift game. Similar to the previous result, the difference between the two
% % approaches is always higher for the prisoner's dilemma class.

% % \begin{figure}[!htbp]
% %   \centering
% %   \includegraphics[width=.33\textwidth]{static/rounds_two_by_two_games_cooperation.pdf}
% %   \includegraphics[width=.33\textwidth]{static/opponents_two_by_two_games_cooperation.pdf}
% %   \includegraphics[width=.33\textwidth]{static/rounds_opponents_two_by_two_games_cooperation.pdf}
% %   \caption{\textbf{Cooperation rates over various social dilemmas for different
% %   limited memory payoffs.} ({\bf A}) If players update based on their last two
% %   rounds with another member of the population. ({\bf B}) If players update
% %   based on their last round with two other members of the population. ({\bf C})
% %   If players update based on their last two rounds with two other members of the
% %   population. We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8,
% %   2.2\}\) across the \(x\) axis, and \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
% %   across the \(y\) axis. Unless explicitly varied, the parameters of the
% %   simulation are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times
% %   10^7$..}
% %   \label{fig:cooperation_rates_two_by_two_part_two}
% % \end{figure}

% \section{Conclusions}\label{section:conclusions}

% Cooperation can be seen at odd, why is it that we choose to help others,
% increasing their payoff, at the expense of decreasing one's own? In spite of all
% the selfish genes’ animal and human communities seem to altruistically help each
% other and cooperate, and evolutionary game theory has helped us shape our
% understanding of the evolution of cooperation.

% Previous evolutionary models often feature a curious inconsistency. When
% modeling how individuals make decisions in each round, these models assume that
% players only remember the last round. However, when modeling how individuals
% update their strategies over time, individuals are assumed to have perfect
% memory.

% Here, we have explored how robust cooperation is as models deviate from the
% perfect memory assumption. Initially we considered the donation game. We showed
% that when the last round payoff is used instead of the expected payoffs,
% cooperation can even evolve when individuals only use a minimum of information,
% however, the evolving cooperation rates are typically lower. The resident
% strategies were both less cooperative and less generous. This effect was only
% intensified we increase the benefit and the strength of selection independently.
% The results showed that as each parameter was being increased the difference
% between the cooperation rates were widening. This indicates that cooperative
% players benefit from being able to interact with everyone in the population.

% We extended our approach and presented results not only based on different
% classes of social games, but also based on different limited memory payoffs. The
% analysis showed that specifically for the case of the prisoner's dilemma games
% the contrast between the two approaches can be significantly large. The
% prisoner's dilemma is one of the most applied types of social games when we
% discuss the evolution of cooperation. Our results indicate that cooperation
% struggles to evolve in the prisoner's dilemma when only a minimum of social
% information is used at the updating stage. Interestingly our simulations also
% showed that in same cases cooperation can benefit from limited memory. This was
% specific for the snowdrift game.

% \section{Acknowledgements}

% This work was supported by the European Research Council Starting Grant 850529:
% E-DIRECT.

% \appendix

% \section{Model Setup}\label{appendix:methods}

% Consider a population of \(N\) individuals where \(N\) is even. At any point in
% time there are at most two different strategies in present in the population.
% More specifically, a mutant strategy played by \(k\) individuals and a resident
% strategy played by \(N - k\) individuals. We assume a pairwise process in which
% strategies spread because they are imitated more often. Each step of the
% evolutionary process consists of two stages; a game stage and an update stage.

% In the game stage, each individual is randomly matched with some other
% individual in the population. Their interaction lasts for a number of turns
% which is not fixed but depends on the continuation probability \(\delta\). At
% each turn the individuals choose between cooperation (\(C\)) and defection
% (\(D\)). Thus, there are four possible outcomes in each turn \(CC, CD, DC\) and
% \(DD\). If both players cooperate they receive the reward payoff \(R\), whereas
% if both players defect they receive the punishment payoff \(P\). If one
% cooperates but the other defects, the defector receives the temptation to
% defect, \(T\), whereas the cooperator receives the sucker's payoff, \(S\). Let
% $\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round, and
% let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector. We present
% results for various values of $\mathcal{U}$ for all the symmetric \(2 \times 2\)
% games.

% A further assumption of our model is that individuals make use of reactive
% strategies when they make decisions in each round. Reactive strategies are a set
% of strategies that take into account only the previous action of the opponent. A
% reactive strategy can be written explicitly as a vector,

% \[s=(y, p, q)\]

% where \(y\) is the probability that the strategy opens with a cooperation and
% \(p, q\) are the probabilities that the strategy cooperates given that the
% opponent cooperated and defected equivalently.

% In the updating stage, two players are randomly drawn from the population, a
% `learner' and a `role model'. The learner adopts the role model's strategy based
% on the Fermi distribution function, %ToDo add reference
